---
title: "COMPSCIX 415.2 Homework 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My Github repository for my assignments can be found at this URL: [https://github.com/larrykubin/compscix-415-2-assignments](https://github.com/larrykubin/compscix-415-2-assignments)

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
```

## Question 1 Answer: Number of observations in the dataset

There are 23,018 observations in the dataset. This is found by using _glimpse_, which presents a summary of the dataset.

```{r}
glimpse(WorldCities)
```

## Question 2 Answer: Show unique regions in the dataset

Below is a list of the unique countries in the first 200 records of the dataset.

```{r}
WorldCitiesHead <- head(WorldCities, 200)
country_col <- WorldCitiesHead$country
unique(country_col)
```

Below is a list of the unique regions in the entire dataset. 

```{r}
region_col <- WorldCities$region
unique(region_col)
```

