---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


## 10.5

### 1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).

In a tibble, all observations have a data typ and the output is limited to 10 rows per page. Also R Studio will show tbl_df instead of data.frame. 

```{r}
print(mtcars)

print(mpg)
```

### 2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?


```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```

### 3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?

You can extract the name of a variable using subsetting as below:

```{r}
df <- tibble(
  x = 1,
  y = 2
)

df$x

```


### 6. What option controls how many additional column names are printed at the footer of a tibble?

You can control this using the width option.

## 12.3.3

### 2. Why does this code fail?

The data is not tidy. The years should not be columns, they should be in the observations list. 

```{r}
table4a %>% 
  gather(1999, 2000, key = "year", value = "cases")

```

```{r}
table4a
```

### 3. Why does spreading this tibble fail? How could you add a new column to fix the problem?


```{r}
people <- tribble(
  ~name,             ~key,    ~value, ~year,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45, 1999,
  "Phillip Woods",   "height",   186, 1999,
  "Phillip Woods",   "age",       50, 2000,
  "Jessica Cordero", "age",       37, 1999,
  "Jessica Cordero", "height",   156, 1999
)

people %>% spread(~name, ~key, ~value)
```


## 12.4.3

### 1. What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra="drop")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
```

### 2. Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?

The remove argument removes input columns from the output data frame. You might set it to FALSE because you want to see the inputs next to the output.


