---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: Larry Kubin
date: March 4, 2018
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The tidyverse packages

### 1. Can you name which package is associated with each task below?

#### Plotting

ggplot2

#### Data munging/wrangling

dplyr

#### Reshaping (speading and gathering) data

tidyr

#### Importing/exporting data

readr

### 2. Now can you name two functions that you’ve used from each package that you listed above for these tasks?

#### Plotting

ggplot(), aes()

#### Data munging/wrangling

select(), filter()

#### Reshaping data

gather(), spread()

#### Importing/exporting data (note that readRDS and saveRDS are base R functions)

read_csv(), read_delim()

## R Basics

### 1. Fix this code with the fewest number of changes possible so it works:

The exclamation point was not a valid character in a variable name. A variable name can only consist of letters, numbers, and the dot and and underscore characters.

```{r}
My_data.name___is.too00ooLong <- c( 1 , 2 , 3 )
```

### 2. Fix this code so it works:

One of the values was missing a quotation mark. Also, vectors are created using the c() function. There was a capital C used to create the function, which caused an error.

```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
```

### 3. Look at the code below and comment on what happened to the values in the vector.

A vector can not have multiple types. Since a string is the most complex type in the vector, all values in the vector are coerced to strings.

```{r}
my_vector <- c(1, 2, '3', '4', 5)
my_vector
```

## Data Import / Export

### Import rail_trail.txt

```{r}
library(tidyverse)

rail_trail <- read_delim("rail_trail.txt", "|")
glimpse(rail_trail)
```


### Export the file into an R-specific format and reload it.

```{r}
saveRDS(rail_trail, file = "rail_trail.rds")

rail_trail_rds = readRDS(file = "rail_trail.rds")

glimpse(rail_trail_rds)
```

## Visualization

### 1. Critique this graphic: give only three examples of what is wrong with this graphic. Be concise.

--

### 2. Reproduce this graphic using the diamonds data set.

--

### 3. The previous graphic is not very useful. We can make it much more useful by changing one thing about it. Make the change and plot it again.



## Data munging and wrangling (6 points)
1. Is this data “tidy”? If yes, leave it alone and go to the next problem. If no, make it tidy. Note: this
data set is called table2 and is available in the tidyverse package. It should be ready for you to use
after you’ve loaded the tidyverse package.
table2
## # A tibble: 12 x 4
## country year type count
## <chr> <int> <chr> <int>
## 1 Afghanistan 1999 cases 745
## 2 Afghanistan 1999 population 19987071
## 3 Afghanistan 2000 cases 2666
## 4 Afghanistan 2000 population 20595360
## 5 Brazil 1999 cases 37737
## 6 Brazil 1999 population 172006362
## 7 Brazil 2000 cases 80488
## 8 Brazil 2000 population 174504898
## 9 China 1999 cases 212258
## 10 China 1999 population 1272915272
## 11 China 2000 cases 213766
5
## 12 China 2000 population 1280428583
2. Create a new column in the diamonds data set called price_per_carat that shows the price of each
diamond per carat (hint: divide). Only show me the code, not the output.
3. For each cut of diamond in the diamonds data set, how many diamonds, and what proportion, have a
price > 10000 and a carat < 1.5? There are several ways to get to an answer, but your solution must
use the data wrangling verbs from the tidyverse in order to get credit.
• Do the results make sense? Why?
• Do we need to be wary of any of these numbers? Why?

## EDA

Take a look at the txhousing data set that is included with the ggplot2 package and answer these questions:
1. During what time period is this data from?
2. How many cities are represented?
3. Which city, month and year had the highest number of sales?
4. What kind of relationship do you think exists between the number of listings and the number of sales?
Check your assumption and show your work.
5. What proportion of sales is missing for each city?
6. Looking at only the cities and months with greater than 500 sales:
• Are the distributions of the median sales price (column name median), when grouped by city,
different? The same? Show your work.
• Any cities that stand out that you’d want to investigate further?
• Why might we want to filter out all cities and months with sales less than 500?