---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: Larry Kubin
date: March 4, 2018
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The tidyverse packages

### 1. Can you name which package is associated with each task below?

#### Plotting

ggplot2

#### Data munging/wrangling

dplyr

#### Reshaping (speading and gathering) data

tidyr

#### Importing/exporting data

readr

### 2. Now can you name two functions that you’ve used from each package that you listed above for these tasks?

#### Plotting

ggplot(), aes()

#### Data munging/wrangling

select(), filter()

#### Reshaping data

gather(), spread()

#### Importing/exporting data (note that readRDS and saveRDS are base R functions)

## R Basics

### 1. Fix this code with the fewest number of changes possible so it works:

The exclamation point was not a valid character in a variable name. A variable name can only consist of letters, numbers, and the dot and and underscore characters.

```{r}
My_data.name___is.too00ooLong <- c( 1 , 2 , 3 )
```

### 2. Fix this code so it works:

One of the values was missing a quotation mark. Also, vectors are created using the c() function. There was a capital C used to create the function, which caused an error.

```{r}
my_string <- c('has', 'an', 'error', 'in', 'it')
```

### 3. Look at the code below and comment on what happened to the values in the vector.

A vector can not have multiple types. Since a string is the most complex type in the vector, all values in the vector are coerced to strings.

```{r}
my_vector <- c(1, 2, '3', '4', 5)
my_vector
```

## Data Import / Export

### Import rail_trail.txt

hello

### Export the file into an R-specific format

here we go
